---
title: "Annotation data for RNAseq/TNSeq of S.pyogenes (2017)"
author: "atb abelew@gmail.com"
date: "`r Sys.Date()`"
output:
 html_document:
  code_download: true
  code_folding: show
  fig_caption: true
  fig_height: 7
  fig_width: 7
  highlight: default
  keep_md: false
  mode: selfcontained
  number_sections: true
  self_contained: true
  theme: readable
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
---

<style>
  body .main-container {
    max-width: 1600px;
}
</style>

```{r options, include=FALSE}
library("hpgltools")
tt <- devtools::load_all("~/hpgltools")
knitr::opts_knit$set(progress=TRUE,
                     verbose=TRUE,
                     width=90,
                     echo=TRUE)
knitr::opts_chunk$set(error=TRUE,
                      fig.width=8,
                      fig.height=8,
                      dpi=96)
old_options <- options(digits=4,
                       stringsAsFactors=FALSE,
                       knitr.duplicate.label="allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size=10))
set.seed(1)
ver <- "20180717"
previous_file <- "index.Rmd"

tmp <- try(sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz"))))

rmd_file <- "01_annotation.Rmd"
```

# Annotation version: `r ver`

This worksheet is intended to gather annotation data for use when looking at
tnseq analyses of S.pyogenes strain 5448.  The analyses were performed in July
2017, but I am re-running them in 2018 to make sure that all the code still works.

There are a few methods of importing annotation data into R.  The following
illustrate some of them.

## Read a genbank accession directly

The function load_genbank_annotations() should download the .gb file from
genbank, read it, and provide a set of data frames which are useful.

```{r gbk2txdb}
mgas_data <- sm(load_genbank_annotations(accession="CP008776"))
genome_size <- GenomicRanges::width(mgas_data$seq)  ## This fails on travis?
mgas_cds <- as.data.frame(mgas_data$cds)
## Get rid of amino acid sequence
mgas_cds <- mgas_cds[-15]
mgas_cds <- mgas_cds[-16]
rownames(mgas_cds) <- mgas_cds[["locus_tag"]]
summary(mgas_data)

nc_features <- as.data.frame(mgas_data$others)
write.csv(nc_features, file="nc_features.csv")
cds_features <- as.data.frame(mgas_data$cds)
write.csv(cds_features, file="cds_features.csv")
```

## Microbesonline

Another source of good annotation data for bacterial species is microbesonline.
Its only real annoyance is that it has a whole new set of identifiers for the
various species it contains, and my function which gathers them apparently
caused it to throw errors internally.  Thus, in order to get the genome ID, I
need to go to microbesonline.org and ask it explicitly; ergo the arbitrary number below.

```{r microbesonline}
##microbe_ids <- get_microbesonline_ids("pyogenes")
##microbe_ids
microbe_ids <- 293653  ## MGAS5005
mgas_df <- load_microbesonline_annotations(microbe_ids)
```

## Read a gff file

In contrast, it is possible to load most annotations of interest directly from the gff files used in
the alignments.  More in-depth information for the human transcriptome may be extracted from biomart.

```{r genome_input, eval=FALSE}
## The old way of getting genome/annotation data
sp_gff <- "reference/mgas_5448.gff"
sp_fasta <- "reference/mgas_5448.fasta"
sp_annotations <- load_gff_annotations(sp_gff, type="gene")
```

# Create expt

We have some annotations and will now create an expt containing the annotations
and count data.

The create_expt() function brings together the annotation data, the count tables
in the sample sheet, and the metadata in the sample sheet into a canonical R
expressionSet.  It then adds a small amount of extra information (sample colors,
for example), and recasts it as an expt.

```{r create_expt}
sp_expt <- sm(create_expt(metadata="sample_sheets/samples_v3.xlsx", gene_info=mgas_cds))
knitr::kable(pData(sp_expt))
```

```{r saveme}
message(paste0("This is hpgltools commit: ", get_git_commit()))
pander::pander(sessionInfo())
this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
message(paste0("Saving to ", this_save))
tt <- sm(saveme(filename=this_save))
```
