---
title: "20190116: Differential Expression for TNSeq of Spyogenes. (2017)"
author: "atb"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
 font-size: 16px
}
</style>

```{r options, include=FALSE}
library("hpgltools")
tt <- devtools::load_all("~/hpgltools")
knitr::opts_knit$set(width=120,
                     progress=TRUE,
                     verbose=TRUE,
                     echo=TRUE)
knitr::opts_chunk$set(error=TRUE,
                      dpi=96)
old_options <- options(digits=4,
                       max.print=120,
                       stringsAsFactors=FALSE,
                       knitr.duplicate.label="allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size=10))
rundate <- format(Sys.Date(), format="%Y%m%d")
previous_file <- "20190115_02_sample_estimation.Rmd"
ver <- "20190115"

tmp <- sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz")))

rmd_file <- "20190115_03_differential_expression.Rmd"
```

# Differential Expression, Spyogenes: `r ver`

This worksheet performs a set of pairwise comparisons of the counts for this tnseq experiment
given the sva surrogate adjustments.

This is being performed via limma/edger/deseq2, which one may rightly argue is not
entirely appropriate for TNSeq data.  On the other hand, as a general metric of
changing fitness, it provides some interesting information.

## Perform Pairwise analyses

```{r initial_estimation, fig.show="hide"}
initial_diff <- sm(all_pairwise(input=rpmi_filt, model_batch="svaseq"))
initial_fsva_diff <- sm(all_pairwise(input=rpmi_filt, model_batch="fsva"))
initial_nobatch_diff <- sm(all_pairwise(input=rpmi_filt, model_batch=FALSE))
```

all_pairwise() is responsible for following:

1.  Creating an experimental model which includes surrogate estimates from svaseq.
2.  Calling limma, EdgeR, and DESEq2 with the data and that model for every
    possible pairwise contrast in the data. (Optionally it will now do EBSeq, too.)
3.  Calling a basic method which simply subtracts log2 scaled values as a form
    of negative control.
4.  Collecting those results.

Among the summaries it provides is a heatmap of how similar each method was for
every contrast.

```{r compare_heatmap}
initial_diff$comparison$heat
## All contrasts are quite similar.
## The lowest correlation coefficient among the methods/contrasts is ~ 0.6 and is
## between the 'real' methods and my stupid basic method, which mostly tells me
## that sva actually did something.
initial_diff$comparison[[7]]
## Here we see that the log2FCs from limma and DESeq2 are very similar.
## The blue line is the identity; so DESeq is seeing slightly higher
## values than limma almost across the board.
initial_diff$comparison[[9]]
## The lower correlation for the basic is clearly coming from that blob near 0 logFC,
## but for the most part, basic agrees with the sva adjusted limma.
```

## Combine the tables

combine_de_tables() does what it says on the tin.  When provided an excel
filename, it will also put the combined tables into the excel file and try to
add in some pretty plots to the resulting excel file.

```{r combine}
keepers <- list("t2t1" = c("rpmi_t2", "thy_t1"),
                "t3t1" = c("rpmi_t3", "thy_t1"),
                "t2zn" = c("rpmi_t2_highzn", "rpmi_t2_lowzn"),
                "t3zn" = c("rpmi_t3_highzn", "rpmi_t3_lowzn"),
                "time_highzn" = c("rpmi_t3_highzn", "rpmi_t2_highzn"),
                "time_lowzn" = c("rpmi_t3_lowzn", "rpmi_t2_lowzn"),
                "t2_rpmi_highzn" = c("rpmi_t2_highzn", "rpmi_t2"),
                "t3_rpmi_highzn" = c("rpmi_t3_highzn", "rpmi_t3"),
                "t2_rpmi_lowzn" = c("rpmi_t2_lowzn", "rpmi_t2"),
                "t3_rpmi_lowzn" = c("rpmi_t3_lowzn", "rpmi_t3")
                )
## Each element of keepers it a name of the contrast to perform followed by a 2 element list
## with the numerator and denominator.
initial_write <- sm(combine_de_tables(initial_diff,
                                      excel=paste0("excel/", rundate, "_initial_diff-v", ver, ".xlsx"),
                                      keepers=keepers))
initial_fsva_write <- sm(combine_de_tables(initial_fsva_diff,
                                           excel=paste0("excel/", rundate, "_initial_diff_fsva-v", ver, ".xlsx"),
                                           keepers=keepers))
initial_nobatch_write <- sm(combine_de_tables(initial_nobatch_diff,
                                              excel=paste0("excel/", rundate, "_initial_diff_nobatch-v", ver, ".xlsx"),
                                              keepers=keepers))

initial_write[["deseq_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_write[["limma_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_write[["deseq_vol_plots"]][["t2_rpmi_highzn"]]$plot
initial_write[["limma_vol_plots"]][["t2_rpmi_highzn"]]$plot

initial_fsva_write[["deseq_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_fsva_write[["limma_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_fsva_write[["deseq_vol_plots"]][["t2_rpmi_highzn"]]$plot
initial_fsva_write[["limma_vol_plots"]][["t2_rpmi_highzn"]]$plot

initial_nobatch_write[["deseq_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_nobatch_write[["limma_vol_plots"]][["t2_rpmi_lowzn"]]$plot
initial_nobatch_write[["deseq_vol_plots"]][["t2_rpmi_highzn"]]$plot
initial_nobatch_write[["limma_vol_plots"]][["t2_rpmi_highzn"]]$plot
```

The data generated by combine_de_tables() is supposed to give me some plots
which we may use to evaluate how well the data came together as well as how
similar the final results are for limma/deseq/edger.

## Find significant genes

```{r significance}
sig_write <- sm(extract_significant_genes(
  initial_write,
  excel=paste0("excel/", rundate, "_significant-v", ver, ".xlsx")))
```

```{r saveme}
message(paste0("This is hpgltools commit: ", get_git_commit()))
pander::pander(sessionInfo())
this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
message(paste0("Saving to ", this_save))
tmp <- sm(saveme(filename=this_save))
```
